Package asserts

### 2020-05-15 evaluation/R/package-asserts.Rmd
Update corpus analysis - revdeps, roxygen, renaming
Add asserts analysis and data

---
title: "Runtime checks"
output: html_document
editor_options: 
  chunk_output_type: console
---

@@

knitr::opts_chunk$set(echo = TRUE)
```

```{r paths, include=FALSE}
source("inc/paths.R")
```

@@

The `r RUNTIME_CHECKS` file contains information about the use of assertions across packages.
To generate it, run the following:

@@

```{r load run}
run <- read_csv(RUNTIME_CHECKS_RUN)
run <- semi_join(run, CORPUS_PACKAGES_DF, by=c("job"="package"))
```

```{r check run packages}
#stopifnot(nrow(run) == nrow(CORPUS_PACKAGES_DF))
```

```{r load runtime checks}
runtime_checks <- read_csv(RUNTIME_CHECKS)
runtime_checks <- semi_join(runtime_checks, CORPUS_PACKAGES_DF, by="package")
```

## Analysis

### By package

```{r}
per_package <- count(runtime_checks, package)
```

```{r}
per_package %>%
  arrange(desc(n)) %>%
  datatable()
```

```{r}
ggplot(per_package, aes(n)) + geom_histogram(binwidth = 1)
```

### By package function

```{r}
per_package_fun <- count(runtime_checks, package, fun_name)
```

```{r}
per_package_fun %>%
  arrange(desc(n)) %>%
  datatable()
```

```{r}
ggplot(per_package_fun, aes(n)) + geom_histogram(binwidth = 1)
```

### By asertion function

```{r}
per_assert <- count(runtime_checks, assert)
```

```{r}
per_assert %>%
  arrange(desc(n)) %>%
  datatable()
```

## Raw data

```{r}
datatable(
  runtime_checks, 
  filter="top", 
  options=list(
    pageLength=50,
    autoWidth=TRUE
  )
)
```

>>>>

---
title: "Package asserts"
output: html_document
editor_options: 
  chunk_output_type: console
---

@@

knitr::opts_chunk$set(echo = TRUE)

source("inc/paths.R")
source("inc/setup.R")
source("inc/latextags.R")

options(repos=CRAN_LOCAL_MIRROR)

create_tags(path(TAGS_DIR, "package-asserts.tex"), prefix="", default=TRUE)
```

@@

The `r CRAN_ASSERTS` file contains information about the use of assertions across packages.
To generate it, run the following:

@@

```{r load corpus}
corpus <- read_csv(CORPUS_DETAILS)
```

```{r load asserts}
cran_asserts <- read_csv(CRAN_ASSERTS)
corpus_asserts <- semi_join(cran_asserts, corpus, by="package")
```

## Typeable and partially typeable

```{r}
TYPE_FUNS <- c(
  "is.array",
  "is.atomic",
  "is.call",
  "is.character",
  "is.complex",
  "is.data.frame",
  "is.data.table",
  "is.double",
  "is.environment",
  "is.factor",
  "is.function",
  "is.integer",
  "is.language",
  "is.list",
  "is.logical",
  "is.matrix",
  "is.name",
  "is.numeric",
  "is.primitive",
  "is.symbol",
  "is.raw",
  "is.vector",

  "is.na",
  "is.null",

  "inherits",
  "is",
  "extends",

  # from desc package
  "is_bool",
  "is_call",
  "is_character",
  "is_formula",
  "is_list",
  "is_logical",
  "is_null",
  "is_string",
  "is_scalar_character",
  "is_scalar_logical",
  "has_no_na",
  "is_string_or_null"
)

AUX_FUNS <- c(
  "stopifnot",
  "|",
  "||",
  "&",
  "&&",
  "!",
  "("
)

remove_type_asserts <- function(x, simplify=FALSE) {
  ret <- if (!is.call(x)) {
    x
  } else {
    fun <- as.character(x[[1]])
    if (fun == "==" || fun == "!=") {
      if (is.call(x[[2]]) && x[[2]][[1]] == "length" && (x[[3]] == 1 || x[[3]] == 1L)) {
        NULL
      } else {
        x
      }
    } else if (any(fun %in% AUX_FUNS)) {
      args <- map(x[-1], remove_type_asserts, simplify=simplify)
      if (is.null(unlist(args))) {
        NULL
      } else if (any(sapply(args, is.null)) && simplify) {
          NULL
      } else {
          x
      }
    } else if (any(fun %in% TYPE_FUNS)) {
      NULL
    } else {
      x
    }
  }

  if (is.list(ret)) {
    ret[[1]]
  } else {
    ret
  }
}

remove_type_asserts(quote(stopifnot(is.integer(x) && is.character2(x))))
remove_type_asserts(quote(stopifnot(is.integer(x) && is.character2(x))), simplify = T)

remove_type_asserts(quote(stopifnot(is.integer(x) | is.character(x))))
remove_type_asserts(quote(stopifnot(length(x) == 1)))
remove_type_asserts(quote(stopifnot(length(x) == 1L)))
```

```{r}
remove_type_asserts_from_args <- function(args, simplify=FALSE) {
  args %>%
  map(~str_c("stopifnot(", ., ")")) %>%
  map(~tryCatch(parse(text=.)[[1]], error=function(e) { message("Unable to parse: ",., " : ", e$message); NA })) %>%
  map(~remove_type_asserts(., simplify=simplify)) %>%
  map(~is.null(.))
}

remove_type_asserts_from_df <- function(df) {
  df %>%
  mutate(
    typed=unlist(remove_type_asserts_from_args(args)),
    partially_typed=unlist(remove_type_asserts_from_args(args, TRUE))
  )
}

# df %>%
#   filter(!partially_typed) %>%
#   select(package, fun_name, args) %>%
#  View()
```

```{r}
summary_asserts <- function(df, prefix) {
  df <- remove_type_asserts_from_df(df)
  typed <- filter(df, typed)
  partially_typed <- filter(df, partially_typed)

  overview_table(
    r(paste(prefix, "asserts"), df),
    r(paste(prefix, "asserts in packages"), count(df, package)),
    r(paste(prefix, "asserts in functions"), count(df, fun_name)),
    r(paste(prefix, "stopifnot ratio"), ratio(filter(df, assert=="stopifnot"), df)),

    r(paste(prefix, "typed asserts"), typed),
    r(paste(prefix, "typed asserts ratio"), ratio(typed, df)),
    r(paste(prefix, "typed asserts packages"), count(typed, package)),
    r(paste(prefix, "typed asserts packages ratio"), ratio(count(typed, package), count(df, package))),
    r(paste(prefix, "typed asserts functions"), count(typed, fun_name)),
    r(paste(prefix, "typed asserts functions ration"), ratio(count(typed, fun_name), count(df, fun_name))),

    r(paste(prefix, "partially typed asserts"), partially_typed),
    r(paste(prefix, "partially typed asserts ratio"), ratio(partially_typed, df)),
    r(paste(prefix, "partially typed asserts packages"), count(partially_typed, package)),
    r(paste(prefix, "partially typed asserts packages ratio"), ratio(count(partially_typed, package), count(df, package))),
    r(paste(prefix, "partially typed asserts functions"), count(partially_typed, fun_name)),
    r(paste(prefix, "partially typed asserts functions ration"), ratio(count(partially_typed, fun_name), count(df, fun_name))),
  )
}
```

```{r summary for cran}
summary_asserts(cran_asserts, "cran")
```

```{r summary for corpus}
summary_asserts(corpus_asserts, "corpus")
```
